What can Issues on GitHub and Google Code tell us about Bug Hazards in Android and Java?

### 2015-02-13 android-stacks.tex
all over
premilinar version to submit

 The high prevalence of NullPointerExceptions and the undocumented runtime exceptions signaled by the Android Platform
 is aligned with the findings of other research works ~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher} and 
~\cite{kechagia2014} respectively - which detected such bug hazards while testing or mining crash reports of Java and Android projects. 
On the one hand, the findings our study emphasize the impact of these bug harzards on the application robustness 
by mining a different source of information: the issues created by developers on GitHub and Google Code. On the other hand, 
our work points to bug hazards that were not detected by previous works (i.e., cross-type wrappings, undocumented checked exceptions and 
undocumented runtime exceptions thrown by third-party libraries) which can also negatively affect the application robustness.

@@

% The Android Application Framework supports an event-driven development model. 
% event driven and inversion of control.
Among the basic elements that comprise an Android application, the Activity 
is the main class of an app (there is usually one
Activity per application window). Each Activity defines a set of hook methods  
(e.g., OnCreate(), onPause()) that are dispatched automatically by the Android 
framework, in response to infrastructure and application events.

@@

By convention, instances of Error represent  unrecoverable conditions which usually results
from failures detected by the Java Virtual Machine due to resource limitations such as OutOfMemoryError;
normally they cannot be handled inside the application.

Instances of RuntimException are implicitly thrown by Java runtime environment when a program violates 
the semantic constraints of the Java programming language (e.g., out-of-bounds array index, divide-by-zero 
error, null pointer references). Some programming languages react to such errors by peremptorily terminating the program; 
other programming languages allow similar approaches e.g.C\#; and others let the program to continue
 its execution in some situations such as the out-of-bounds array index e.g., C++. 
The programs are not expected to handle such runtime exceptions signaled by the runtime environment according 
to the Java Specification ~\cite{gosling2000java}.

The user-defined exceptions can be either checked or runtime. There is a long-lasting debate 
about the pros and cons of both approaches~\cite{javatut,stackoverlow,debate}
%\footnote{152 questions in Stackoverflow are related to this debate}. 
Section ~\ref{sec:best} presents
a set of best practices related to each exception type.

@@

Based on the assumption that regardless the issue lables, every exception stack
trace contains relevant information concerning the exception structure of the
projects analyzed - and therefore can review bug hazards on the exception-related code -  
we opted for not restricting the analysis only on defect issues\footnote{We conducted the same analysis 
on the defect issues only and the top exceptions were quite similar.
 Due to space limitation we limit to present the general analysis here. More detailed analysis
on the defect issues can be found at:\url{https://github.com/souzacoelho/exceptionminer}}.

\subsection{Android Apps in Google Code}
Google Code contains largely used open-source Android apps (e.g. K9Mail \footnote{K9Mail moved to Github but as a way of not loosing the project history 
it advises their users to report bugs on Google Code issue tracker: https://github.com/k9mail/k-9/wiki/LoggingErrors.}),
 however, differently from GitHub, Google Code does not provide an API to access the information related
 to hosted projects \footnote{Google code used to provided a Web service to such repository, but it was deactivated in June 2013 what Google called a `clean-up action".}.
To overcome this limitation we needed to implement a Web Crawler (incorporated in ExceptionMiner tool described next) that navigated 
through the web interface of Google Code projects extracting all issues and issue comments and storing in a relational database for later analysis.

@@

\emph{\textbf{Exception Stack Trace Distiller.}}
This component distills the information that composes a stack trace.
 Some of the attributes extracted from the stack trace are:
 the root exception and its signaler, the exception wrappers and their corresponding signalers. 
This component also distills fine grained information of each attribute such as: the classes and packages associated to them.
This component is based on a combination between a parser (based on regular expressions) 
and heuristics able to identify and filter exception names and stack traces inline with text. 
In contrast to existing issue parsing solutions such as Infozilla, the parser
created in this work can discover stack traces mixed with log information\footnote{In several 
exception stack traces, the exception frames were preceeded by logging information e.g., 
"03-01 15:55:01.609 (7924): at android.app.ActivityThread.access\$600(ActivityThread.java:127) 
"which could not be detected by existing tools.)}.

@@

\begin{table}
\centering
\begin{tabular}{lrr}
\hline
\bfseries{Category} & \bfseries{\% Occurrences on stacks} \\
\hline
Programming logic (java.lang and util) & 52,0\%\\
Resources (IO, Memory, Batery) & 23,9\% \\
Security & 4,1\%\\
Concurrency & 2,9\% \\
Backward compatibility & 5,5\% \\
Specific Exceptions & 4,9\%\\
General (Error, Exception, Runtime) & 6,7\%\\
\end{tabular}
\caption{Identifying the concerns related to root exceptions}
\label{tab:tophundrend}
\end{table}

@@

\emph{\textbf{Runtime Exception wrapping Error.}} From Table~\ref{tab:wrappingandroid}, we could observe that most of such wrappings
were performed by Android platform (50.7\%). Inspecting each of them revealed that the android.os.AsyncTask, was responsible for 44\% 
of these wrappings. This class is responsible for dealing with asynchronous tasks in Android.
The code snippet below shows a general catch clause that converts any instance of Throwable (signaled during the execution
of an asynchronous task) into an instance of RuntimeException and re-throw it.  
Table~\ref{tab:exampeswrap} exemplifies some of the wrappings performed in this code snipet:  
an instance of RuntimeExceptions wrapping instances of OutOfMemoryError and StackOverflowError. 
Such wraps mask an unrecoverable error wrapping on a general runtime exception.

Moreover, such wrapping also pointed out to another best practice violation: the methods from Android platform
signaling direct instances of RuntimeException. Using such general exception, is considered a bad practice 
according to Java specification and compiled guidelines, because it loses contextual information about the exception.

@@

@@

\emph{\textbf{Runtime wrapping Checked Exception.}} This wrapping was responsible for 49.5\% of
the cross-type wrappings.  From this set 50\%  were performed on methods defined on Android platform. 
We could observe that approximately 41\% of them were performed by a single class,
 named ActivityThreat. This class is responsible for managing the life cycle of  Android Activities - 
invoking the life cycle methods (i.e. onCreate(), onPause()) for every Activity 
(window) of an Android app. This class wraps every instance of java.lang.Exception 
signaled withing the life cycle methods in a general RuntimeException and re-throws. 
We observed that it was a common design decision in other methods of Android platform
(although usually considered a bad practice).

@@

\subsection{On the need of rescues to the almost infeasible task of preventing uncaught exceptions}
In this study we could observe undocumented runtime exceptions thrown by third party code,
and even undocumented checked exception thrown by a JNI interface.
Such undocumented exceptions make difficult, and most of the times infeasible
for the client code to protect against `unforeseen" situations that may happen 
while calling a library code.

One may think that the solution for the uncaught exceptions may be to define a general handler, 
which is responsible for handling any exception that is not
adequately handled inside the applications. Although this 
solution may prevent  the system of abruptly crashing,
 such general handler will not have enough
contextual information to adequately handle the exception, 
a possible handler task is: to present a message to the user
 and restart the application. However, such handler cannot replace a carefully designed exception 
handling policy ~\cite{Robil00}, which can hardly be designed in the absence of 
third-party documentation on the exceptions that
may the thrown by it. On the other hand, documenting runtime exceptions is a tedious and error prone task, to help developers
mitigate this problem tools should be developed to automate the documentation of runtime exceptions
scraping from library code, few solutions in this directions have been proposed so far ~\cite{van2005combining}.

\subsection{On the null pointer problem}
The null references was firstly introduced by Tony Hoare in ALGOL W, which after some years he called his ``one-billion-dollar mistake"~\cite{hoare2}.

%:
%\emph{``I call it my billion-dollar mistake. It was the invention of the null reference in 1965 [...] 
%This has led to innumerable errors, vulnerabilities, and system 
%crashes, which have probably caused a billion dollars of pain and damage in the last forty years."}

In this sudy, the null references were, infact, responsible for several reported issues - providing more evidences to Hoare's speech.
This observation emphasizes the importance of solutions to avoid NullPointerExceptions, such as:
(i) lightweight intra-method null pointer analysis supported by Java8 @Nullable annotations~\footnote{They are used by tools e.g., Eclipse, IntelliJ, Android Studio 0.5.5 (release Apr. 2014) to detect potential 
null pointer dereferences at compile time.};
(ii) inter-method null pointer analysis tools such as the one proposed by Nanda and Sinha ~\cite{nanda2009accurate};
or (iii) language designs which avoid null pointers, such 
as Monads ~\cite{Walde95} (i.e., used in functional languages for values that may not be available 
or computations that may fail) could improve the robustness of Java programs.

\subsection{The exception handling confusion problem}
When (mis)applied, the exception wrapping can make the exception-related code
 more complex and lead to what we call an \emph{exception handling confusion problem}.
Such problem can lead the program to an unpredictable state in the presence of exceptions,
as illustrated by the scenario on which a checked exception wrapped an OutOfMemoryError. 
Currently there is no way of enforcing Java exception type conventions during program development.
Hence, further investigation is needed on finding ways to help developers in dealing with such
 problem, either preventing odd wrappings or enabling the developer to better deal with them or even
empirical studies on the real usefulness of Java hybrid exception model.

@@

@@

%\noindent
\emph{Internal Validity.} We used a heuristics-based parser to mine
exceptions from issues.  Our parsing strategy was conservative by default; for
example, we only considered exception names using a fully qualified class name
as valid exception identifiers, while, in many cases, developers use the
exception name in issue description. Conservative parsing may minimize false
positives, which was our initial target, but also tends to increase false
negatives, which means that some cases may have not been identified as
exceptions or stack traces. Our limited manual inspection did not reveal such
cases. Moreover, in this study we manually classified the exceptions 
related to programming mistakes. To ensure the quality of the analysis, 
we calculated the interrater agreement after three independent 
developers classified a randomly selected sample (of 25 exception 
types from the total of 100); the interater agreement was high (96\%). 
Another threat relates to the fact that parts of our analysis 
are based on the availability of stack traces on issues reported on Github and Googlecode projects. 
In using these datasets, we make an underlying assumption: the stack traces reported on issues are 
representative of valid crash information of the applications. 
One way to mitigate such threat would be to access to the full 
set of crash data per application. Although some services exist 
to collect crash data from mobile applications ~\cite{BugSe14,BugSn14,Googl14,Acra14},
they do not provide open access to the crash reports of their client applications.
In our study, we mitigated this threat by manually inspecting
the source code associated to a subset of the reported exception stack traces.
Such subset comprises the stack traces related to the main findings 
of the study (e.g., ``undocumented runtime and checked exceptions",
and cross-type wrappings").
%\noindent

@@

\textit{Analysis and Use of Stack Trace Information.} Several research works have
investigated the use of stack trace information to support bug classification
and clustering~\cite{wang2013improving, kim2011crash, dhaliwal2011classifying},
fault-proneness prediction models~\cite{kim2013predicting} and even automated
bug fixing tools~\cite{sinha2009fault}. Kim et al.~\cite{kim2011crash} use an
aggregated form of multiple stack traces available in crash reports to detect
duplicate crash reports and to predict if a given crash will be fixed. Dhaliwal
et al.~\cite{dhaliwal2011classifying} proposed a crash grouping approach that
can reduce bug fixing time in approximately 5\%. Wang et
al.~\cite{wang2013improving} propose an approach to identify correlated crash
types and describe a fault localization method to locate and rank files related
to the bug described on a stack trace. Schroter et al.~\cite{schroter2010stack}
conducted an empirical study on the usefulness of stack traces for bug fixing
and showed that developers fixed the bugs faster when failing stack traces were
included on bug issues.  In a similar study, Bettenburg et
al.~\cite{bettenburg2008makes} identify stack traces as the second most stack
trace feature for developers.  Sinha et al.~\cite{sinha2009fault} proposed an
approach that uses stack traces to guide a dataflow analysis for locating and
repairing faults that are caused by the implicitly signaled exceptions. Kim
at al.~\cite{kim2013predicting} proposed an approach to predict the
crash-proneness of methods based information extracted from stack traces and
methods' bytecode operations.  They observed that most of the stack traces were
related to NullPointerException and other implicitly thrown exceptions had
the higher prevalence on the analyzed set of stacks. Kechagia and Spinellis ~\cite{kechagia2014}
examined the stack traces embedded on crash reports sent by 1,800 Android apps 
to a crash report management service (i.e., BugSense). They found that 19% of such stack traces
were caused by unchecked and undocumented exceptions thrown by methods defined on 
Android API (level 15). Our work differs from Kechagia and Spinellis since it is not based on
stack traces mined from issues reported by open source developers on GitHub and Googlecode.  
Moreover, our study mapped the origin of each exception 
(i.e., libraries, the Android platform or the application itself) and investigated
the adoption of best practices based on the analysis of stack trace information.
Our work also identified the type of each exception mined from issues
(classifying them as Error, Runtime or Checked) based on the source code
analysis of the exception hierarchy and analized the exception wrappings that can 
happen during the exception propagation.  Such analysis revealed intrigging 
bug hazards such as the cross-type exception wrappings not discussed in previous works.

\textit{Empirical Studies on Exception Handling Defects.} 
Cabral and Marques~\cite{cabral2007exception} analyzed the
source code of 32 open-source systems, both for Java and .NET. They
observed that the actions inside handlers were very simple (e.g., logging and present a
message to the user). Coelho et al.  ~\cite{coelho2008assessing} performed an 
empirical study considering the fault-proneness of aspect-oriented implementations 
for handling exceptions. Two releases of both Java and AspectJ implementation of a 
product line were assessed as part of that study. Based on the use of an exception
flow analysis tool, the study revealed that the AOP  refactoring increased the 
number of uncaught exceptions, degrading the robustness of the AO version of every analyzed system.
The main limitation of approaches based static analysis based approaches are the number of false
positives they can generate, and the problems the faced when dealing with reflection libraries 
and dynamic class loading. Pingyu and Elbaum~\cite{Zhang12} were the first to perform
an empirical investigation of issues, related to exception-related bugs, on Android projects.  
They perform a small scale study on which they manually inspected the issues of 
5 Android applications. They observed that 29\% had to do with poor
exceptional handling code, this empirical study was used to motivate the development of a tool
aiming at amplifying existing tests to validate exception 
handling code associated with external resources. This work inspired ours,
 which automaticaly mined the exception stack traces embedded on issues 
reported on 639 open source Android projects. The goal of our study was
to identify common bug hazards on the exception related code that can lead to 
failures such as uncaught exceptions.

@@

\textit{Empirical studies using Android apps.} Ruiz et al.~\cite{Ruiz12}
investigated the degree of reuse across applications in Android Market, the
study showed that almost 23\% of the classes inherited from a base class in the
Android API, and that 217 mobile apps were reused completely by another mobile
app. Pathak et al.~\cite{Patha11} analyzed bug reports and developers
discussions of Android platform and found out that approximately 20\% of
energy-related bugs in Android occurred after an OS update. McDonnell et
al.~\cite{McDon13} conducted a case study of the co-evolution behavior of
Android API and 10 dependent applications using the version history data found
in github. The study found that approximately 25\% of all methods in the client
code used the Android API, and that the methods reusing fast-evolving APIs were
more defect prone then others. Vásquez et al.~\cite{Linar13} analyzed
approximately 7K free Android apps and observed that the last successful apps
used Android APIs that were on average 300\% more change-prone than the APIs
used by the most successful apps. Pingyu and Elbaum~\cite{Zhang12} analyzed bug
reports of 5 Android applications an observed that 29\% had to do with poor
exceptional handling code. Our work differs from the others as it aims at
distilling the information of bug reports distilling the uncaught exceptions 
occurring at mobile applications in Github and Google Code. Such analysis discusses about
the causes of bugs related to exceptions across applications available and what the
characteristics of the stack traces can tell us about the exception structure of
those applications.

>>>>

 The high prevalence of NullPointerExceptions is aligned with the findings of other research works ~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher} as well as the undocumented runtime 
exceptions signaled by the Android Platform ~\cite{kechagia2014}. If on the one hand, the findings our
 study emphasize the impact of these bug harzards on the application robustness 
by mining a different source of information: the issues created by developers on GitHub and Google Code. On the other hand, 
our work points to bug hazards that were not detected by previous works (i.e., cross-type wrappings, undocumented checked exceptions and 
undocumented runtime exceptions thrown by third-party libraries) which can also negatively affect the application robustness.

@@

% The Android Application Framework supports an event-driven development model. 
% event driven and inversion of control.
%Among the basic elements that comprise an Android application, the Activity 
%is the main class of an app (there is usually one
%Activity per application window). Each Activity defines a set of hook methods  
%(e.g., OnCreate(), onPause()) that are dispatched automatically by the Android 
%framework, in response to infrastructure and application events.

@@

By convention, instances of Error represent  unrecoverable conditions which usually results
from failures detected by the Java Virtual Machine due to resource limitations such as OutOfMemoryError;
normally they cannot be handled inside the application.  Instances of RuntimException are implicitly 
thrown by Java runtime environment when a program violates 
the semantic constraints of the Java programming language (e.g., out-of-bounds array index, divide-by-zero 
error, null pointer references). Some programming languages react to such errors by peremptorily terminating the program; 
other programming languages allow similar approaches e.g.C\#; and others let the program to continue
 its execution in some situations such as the out-of-bounds array index e.g., C++. 
The programs are not expected to handle such runtime exceptions signaled by the runtime environment according 
to the Java Specification ~\cite{gosling2000java}. The user-defined exceptions can be either checked
 or runtime. There is a long-lasting debate 
about the pros and cons of both approaches~\cite{javatut,stackoverlow,debate}
%\footnote{152 questions in Stackoverflow are related to this debate}. 
Section ~\ref{sec:best} presents
a set of best practices related to each exception type.

@@

Based on the assumption that regardless the issue lables, every exception stack
trace contains relevant information concerning the exception structure of the
projects analyzed - and therefore can review bug hazards on the exception-related code -  
we opted for not restricting the analysis only on defect issues.

%\footnote{We conducted the same analysis 
%on the defect issues only and the top exceptions were quite similar.
% Due to space limitation we limit to present the general analysis here. More detailed analysis
%on the defect issues can be found at:\url{https://github.com/souzacoelho/exceptionminer}}.

\subsection{Android Apps in Google Code}
Google Code contains largely used open-source Android apps (e.g. K9Mail\footnote{K9Mail moved to Github but as a way of not loosing the project history it advises their users to report bugs on Google Code issue tracker: https://github.com/k9mail/k-9/wiki/LoggingErrors.}),
 however, differently from GitHub, Google Code does not provide an API to access the information related
 to hosted projects\footnote{Google code used to provided a Web service to such repository, but it was deactivated in June 2013 what Google called a `clean-up action".}.
To overcome this limitation we needed to implement a Web Crawler (incorporated in ExceptionMiner tool described next) that navigated 
through the web interface of Google Code projects extracting all issues and issue comments and storing in a relational database for later analysis.

@@

\emph{\textbf{Exception Stack Trace Distiller.}}
This component distills the information that composes a stack trace.
 Some of the attributes extracted from the stack trace are:
 the root exception and its signaler, the exception wrappers and their corresponding signalers. 
This component also distills fine grained information of each attribute such as: the classes and packages associated to them.
This component is based on a combination between a parser (based on regular expressions) 
and heuristics able to identify and filter exception names and stack traces inline with text. 
In contrast to existing issue parsing solutions such as Infozilla, the parser
created in this work can discover stack traces mixed with log information\footnote{In several 
exception stack traces, the exception frames were preceeded by logging information e.g., 
"03-01 15:55:01.609 (7924): at android.app.ActivityThread.access\$600(ActivityThread.java:127) 
"which could not be detected by existing tools.}.

@@

\begin{table}
\centering
\begin{tabular}{lrr}
\hline
\bfseries{Concern} & \bfseries{\% Occurrences on stacks} \\
\hline
Programming logic (java.lang and util) & 52,0\%\\
Resources (IO, Memory, Batery) & 23,9\% \\
Security & 4,1\%\\
Concurrency & 2,9\% \\
Backward compatibility & 5,5\% \\
Specific Exceptions & 4,9\%\\
General (Error, Exception, Runtime) & 6,7\%\\
\hline
\end{tabular}
\caption{Identifying the concerns related to root exceptions}
\label{tab:tophundrend}
\end{table}

@@

\emph{\textbf{Runtime Exception wrapping Error.}} From Table~\ref{tab:wrappingandroid}, we could observe that most of such wrappings
were performed by Android platform (50.7\%).
The code snippet below was extracted from Android and shows a general catch clause 
that converts any instance of Throwable (signaled during the execution
of an asynchronous task) into an instance of RuntimeException and re-throws it.  
Table~\ref{tab:exampeswrap} presents exceptions wrapped in this code snipet.  
Such wrappings mask an unrecoverable Error on a general runtime exception.

@@

\emph{\textbf{Runtime wrapping Checked Exception.}} This wrapping was responsible for 49.5\% of
the cross-type wrappings.  From this set 50\%  were performed on methods defined on Android platform. 
We observed that it was a common implementation practice in the methods of Android platform.
However, using such general exception, is considered a bad practice according to Java specification 
and compiled guidelines, because it loses contextual information about the exception.

@@

@@

\emph{\textbf{The exception handling confusion problem.}}
When (mis)applied, the exception wrapping can make the exception-related code
 more complex and lead to what we call an \emph{exception handling confusion problem}.
Such problem can lead the program to an unpredictable state in the presence of exceptions,
as illustrated by the scenario on which a checked exception wrapped an OutOfMemoryError. 
Currently there is no way of enforcing Java exception type conventions during program development.
Hence, further investigation is needed on finding ways to help developers in dealing with such
 problem, either preventing odd wrappings or enabling the developer to better deal with them or even
empirical studies on the real usefulness of Java hybrid exception model.

@@

\emph{\textbf{On the null pointer problem.}}
The null references was firstly introduced by Tony Hoare in ALGOL W, which after some years he called his ``one-billion-dollar mistake"~\cite{hoare2}.
In this sudy, the null references were, infact, responsible for several reported issues - providing more evidences to Hoare's speech.
This observation emphasizes the importance of solutions to avoid NullPointerExceptions, such as:
(i) lightweight intra-method null pointer analysis supported by Java8 @Nullable annotations~\footnote{They are used by tools e.g., Eclipse, IntelliJ, Android Studio 0.5.5 (release Apr. 2014) to detect potential 
null pointer dereferences at compile time.};
(ii) inter-method null pointer analysis tools such as the one proposed by Nanda and Sinha ~\cite{nanda2009accurate};
or (iii) language designs which avoid null pointers, such 
as Monads ~\cite{Walde95} (i.e., used in functional languages for values that may not be available 
or computations that may fail) could improve the robustness of Java programs.

\emph{\textbf{On the need of rescues to the almost infeasible task of preventing uncaught exceptions}}
In this study we could observe undocumented runtime exceptions thrown by third party code,
and even undocumented checked exception thrown by a JNI interface.
Such undocumented exceptions make difficult, and most of the times infeasible
for the client code to protect against `unforeseen" situations that may happen 
while calling a library code. One may think that the solution for the uncaught exceptions may be to define a general handler, 
which is responsible for handling any exception that is not
adequately handled inside the applications. Although this 
solution may prevent  the system of abruptly crashing,
 such general handler will not have enough
contextual information to adequately handle the exception, 
a possible handler task is: to present a message to the user
 and restart the application. However, such handler cannot replace a carefully designed exception 
handling policy ~\cite{Robil00}, which can hardly be designed in the absence of 
third-party documentation on the exceptions that
may the thrown by it. On the other hand, documenting runtime exceptions is a tedious and error prone task, to help developers
mitigate this problem tools should be developed to automate the documentation of runtime exceptions
scraping from library code, few solutions in this directions have been proposed so far ~\cite{van2005combining}.

@@

%\noindent
\emph{Internal Validity.} We used a heuristics-based parser to mine
exceptions from issues.  Our parsing strategy was conservative by default; for
example, we only considered exception names using a fully qualified class name
as valid exception identifiers, while, in many cases, developers use the
exception name in issue description. Conservative parsing may minimize false
positives, which was our initial target, but also tends to increase false
negatives, which means that some cases may have not been identified as
exceptions or stack traces. Our limited manual inspection did not reveal such
cases. Moreover, in this study we manually mapped the concerns related to exceptions.
 To ensure the quality of the analysis, we calculated the interrater agreement after three independent 
developers classified a randomly selected sample (of 25 exception 
types from the total of 100); the interater agreement was high (96\%). 
Another threat relates to the fact that parts of our analysis 
are based on the availability of stack traces on issues reported on Github and Googlecode projects. 
In using these datasets, we make an underlying assumption: the stack traces reported on issues are 
representative of valid crash information of the applications. 
One way to mitigate such threat would be to access to the full 
set of crash data per application. Although some services exist 
to collect crash data from mobile applications ~\cite{BugSe14,BugSn14,Googl14,Acra14},
they do not provide open access to the crash reports of their client applications.
In our study, we mitigated this threat by manually inspecting
the source code associated to a subset of the reported exception stack traces.
Such subset comprises the stack traces related to the main findings 
of the study (e.g., ``undocumented runtime and checked exceptions",
and cross-type wrappings").
%\noindent

@@

\textit{Analysis and Use of Stack Trace Information.} Several research works have
investigated the use of stack trace information to support: bug classification
and clustering~\cite{wang2013improving, kim2011crash, dhaliwal2011classifying},
fault prediction models~\cite{kim2013predicting}, automated
bug fixing tools~\cite{sinha2009fault} and also the analysis of Android APIs~\cite{kechagia2014}. 
Kim et al.~\cite{kim2011crash} use an
aggregated form of multiple stack traces available in crash reports to detect
duplicate crash reports and to predict if a given crash will be fixed. Dhaliwal
et al.~\cite{dhaliwal2011classifying} proposed a crash grouping approach that
can reduce bug fixing time in approximately 5\%. Wang et
al.~\cite{wang2013improving} propose an approach to identify correlated crash
types and describe a fault localization method to locate and rank files related
to the bug described on a stack trace. Schroter et al.~\cite{schroter2010stack}
conducted an empirical study on the usefulness of stack traces for bug fixing
and showed that developers fixed the bugs faster when failing stack traces were
included on bug issues.  In a similar study, Bettenburg et
al.~\cite{bettenburg2008makes} identify stack traces as the second most stack
trace feature for developers.  Sinha et al.~\cite{sinha2009fault} proposed an
approach that uses stack traces to guide a dataflow analysis for locating and
repairing faults that are caused by the implicitly signaled exceptions. Kim
at al.~\cite{kim2013predicting} proposed an approach to predict the
crash-proneness of methods based information extracted from stack traces and
methods' bytecode operations.  They observed that most of the stack traces were
related to NullPointerException and other implicitly thrown exceptions had
the higher prevalence on the analyzed set of stacks. Kechagia and Spinellis~\cite{kechagia2014}
examined the stack traces embedded on crash reports sent by 1,800 Android apps 
to a crash report management service (i.e., BugSense). They found that 19% of such stack traces
were caused by unchecked and undocumented exceptions thrown by methods defined on 
Android API (level 15). Our work differs from Kechagia and Spinellis since it is not based on
stack traces mined from issues reported by open source developers on GitHub and Googlecode.  
Moreover, our study mapped the origin of each exception 
(i.e., libraries, the Android platform or the application itself) and investigated
the adoption of best practices based on the analysis of stack trace information.
Our work also identified the type of each exception mined from issues
(classifying them as Error, Runtime or Checked) based on the source code
analysis of the exception hierarchy and analized the exception wrappings that can 
happen during the exception propagation.  Such analysis revealed intrigging 
bug hazards such as the cross-type exception wrappings not discussed in previous works.

\textit{Empirical Studies on Exception Handling Defects.} 
Cabral and Marques~\cite{cabral2007exception} analyzed the
source code of 32 open-source systems, both for Java and .NET. They
observed that the actions inside handlers were very simple (e.g., logging and present a
message to the user). Coelho et al.  ~\cite{coelho2008assessing} performed an 
empirical study considering the fault-proneness of aspect-oriented implementations 
for handling exceptions. Two releases of both Java and AspectJ implementations were 
assessed as part of that study. Based on the use of an exception
flow analysis tool, the study revealed that the AOP  refactoring increased the 
number of uncaught exceptions, degrading the robustness of the AO version of every analyzed system.
The main limitation of approaches based static analysis based approaches are the number of false
positives they can generate, and the problems the faced when dealing with reflection libraries 
and dynamic class loading. Pingyu and Elbaum~\cite{Zhang12} were the first to perform
an empirical investigation of issues, related to exception-related bugs, on Android projects.  
They perform a small scale study on which they manually inspected the issues of 
5 Android applications. They observed that 29\% had to do with poor
exceptional handling code, this empirical study was used to motivate the development of a tool
aiming at amplifying existing tests to validate exception 
handling code associated with external resources. This work inspired ours,
 which automaticaly mined the exception stack traces embedded on issues 
reported on 639 open source Android projects. The goal of our study was
to identify common bug hazards on the exception related code that can lead to 
failures such as uncaught exceptions.

@@

\textit{Empirical studies using Android apps.} Ruiz et al.~\cite{Ruiz12}
investigated the degree of reuse across applications in Android Market, the
study showed that almost 23\% of the classes inherited from a base class in the
Android API, and that 217 mobile apps were reused completely by another mobile
app. Pathak et al.~\cite{Patha11} analyzed bug reports and developers
discussions of Android platform and found out that approximately 20\% of
energy-related bugs in Android occurred after an OS update. McDonnell et
al.~\cite{McDon13} conducted a case study of the co-evolution behavior of
Android API and 10 dependent applications using the version history data found
in GitHub. The study found that approximately 25\% of all methods in the client
code used the Android API, and that the methods reusing fast-evolving APIs were
more defect prone then others. Vásquez et al.~\cite{Linar13} analyzed
approximately 7K free Android apps and observed that the last successful apps
used Android APIs that were on average 300\% more change-prone than the APIs
used by the most successful apps. Our work differs from the others as it aims at
distilling stack trace information of bug reports and combine such information
with bytecode analysis, source code analysis and manual inspections
to identify bug hazards on the exception handling code of Android apps.
